Improving \textsl{Gaia} parallaxes with a data-driven model \\ of the color--magnitude diagram

### 2017-05-02 paperDistance/paper.tex
paper tweaks ;

In this paper we use the noise deconvolved observed \cmd\ as a prior to infer more precise parallax pdfs from \gaia\ measurements.
Here we will demonstrate this method with a simpler toy model. Instead of using the 2D distribution of the \cmd\ as our prior, for this simpler model we use a 2D distribution that is a 1D Gaussian in the y-direction, with a running mean $\mu = mx + b$ and some thickness $t$, as our prior. The $1\sigma$ contours as well as (noise free) samples from the distribution are shown in Figure \ref{fig:toy}, upper left panel.
Noisy samples from this distribution are shown in the upper right panel.
Using a representation of this distribution as our prior we can infer some posterior belief of $y_{\true,n}$, a true sample of the distribution, from $y_n$, a noisy sample of the distribution.
Here the thickness of the distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
To set up the problem, we have the true underlying distribution of points and the noisy data

\begin{eqnarray}
y_{\true, n} \, &=& \, m_{\true}\,x_n + b_{\true} + \Delta_n \\
y_n \, &=& \, y_{true,n} + \epsilon_n
\label{eq:ytrue}
\end{eqnarray}

where $\Delta_n$ is drawn from the Gaussian function $\mathcal{N}(\Delta_n \given 0, t^2)$, and $\epsilon_n$ is drawn from the Gaussian function $\mathcal{N}(\epsilon_n \given 0, \sigma_n^2)$, so $\sigma_n$ is the measurement uncertainty.
These represent true and noisy draws from a distribution that is a straight line with some thickness $t$, and are visualized in the upper panels of Figure \ref{fig:toy}

First we determine the underlying distribution that the noisy data $y_n$ are drawn from. The $1\sigma$ contours of this noise deconvolved distribution are shown as the blue lines in the lower left panel of Figure \ref{fig:toy}.
This noise deconvolved distribution becomes our prior.
We then use this pior to infer a posterior belief about the true values, $p(y_{\true,n} \given y_n, \tilde{\sigma}_n^2)$, whose expectation value and $1\sigma$ uncertainty is shown as the blue points in the lower panels of Figure \ref{fig:toy}.
When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance $t^2$ and the measurement variance $\sigma_n^2$.
Figure \ref{fig:toy}, upper right, shows 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points. Here we highlight 5 random points in black to highlight the method, but the same method is being applied to all the data.
We learn the noise-deconvolved distribution using marginalized maximum likelihood estimation (MML), which marginalizes over the true data values (we never need to know them) and maximizes the likelihood of the data. Here our likelihood for each individual datum is

@@

Here is a good point to step back and acknowledge that our method is odd/strange/different/new (which maybe you're also currently thinking but can't quite express why yet). We are using the set of data to determine the noise-deconvolved distribution, and then using that noise-deconvolved distribution to make inferences on the data. We are technically using the data twice! If we were going to be fully, technically correct, we should learn the distribution, but leave one datum out; the one datum we want to infer the true y value for\footnote{DWH: Remember what this brain dump is for}. Leaving one out, and then estimating the prior and then inferring the posterior for each data point is way too expensive for the 1024 points in our toy model, and also for the 1.2 million stars in our actual data set. Luckily with large data sets, even at 1024, 1 datum is $\sim0.1\%$ of your dataset and therefore leaving that point out will not change the MML parameters of your noise-deconvolved distribution, and with the \gaia\ dataset of over a million stars, we can safely step into the Empirical Bayes regime.

@@

Since everything is Gaussian, this can be done analytically by multiplying Gaussians together. The blue points in Figure \ref{fig:toy}, lower right panel, represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$ for the highlighted black points. Finally, Figure \ref{fig:toy}, lower right panel, shows the full set of posterior beliefs for all $y_n$. The points are colored by their measurement variance $\sigma_n^2$ to show that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior, while more certain measurements remain closer to their measured values $y_n$. By doing this we have better estimates of the true y value for each point. Bringing the toy model back to our original problem, using the noise-deconvolved CMD as our prior allows us to get better estimates of the parallax for each star.

@@

We make the following assumptions. The method we present is correct
and justifiable under these assumptions, each of which is questionable
in its own right. We return to criticize these assumptions---and the
method that flows from them---in the Discussion below.
\begin{itemize}
\item We make an assumption of \emph{stationarity}; that is, that the
  stars measured at lower signal-to-noise have analogs measured at
  higher signal-to-noise.  In detail, this assumption is wrong, of
  course (for example, distant halo stars are different than local
  disk stars), but the stationarity assumption is fairly weak; as long
  as there is support in the high signal-to-noise stars for the types
  of stars seen at low signal-to-noise, the results will not be
  strongly biased, but this assumption is deep and fundamental to this
  work.
\item We assume that we have large numbers of stars, large enough that
  an empirical-Bayes or maximum-marginalized likelihood is a safe
  approximation to full Bayesian inference. This assumption will be
  least true in the least-populated parts of the \cmd. In the extreme
  case of a stellar type with only one (statistically distinct)
  exemplar in the data set (e.g. a white dwarf), this assumption will be maximally wrong.
\item We assume that the \tgas\ parallax uncertainties dominate the
  noise in any estimates of absolute magnitude. We further assume that
  the parallax and color uncertainties are Gaussian in form with
  correctly known variances. In detail, we assume that the [INSERT
    CORRECT TERMINOLOGY FROM GAIA HERE] parallax uncertainty estimates
  are correct. Here we are making use of the fact that, for
  de-noising, under-estimation of measurement noise is conservative
  (it leads to under-deconvolution).
\item We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
  as correct in their median dust estimates, and their
  effects on color and magnitude. As we optimize the empirical-Bayes model, we iteratively update each star's
  dust correction as its updated, inferred distance. Our assumption is that these corrections are
  good enough, and uncertainties in these are not dominant, for either
  color or absolute magnitude.
\item We assume that the \cmd\ can be represented with a mixture of
  Gaussians in a particular transformed space. This mixture we fix
  (with only heuristic analysis) at 128, 512, and 2048 Gaussians in
  different experiments.
\item We make no use of stellar models; our only assumptions about
  stellar physics are generic and implicit (for example, that the
  \cmd\ is somehow smooth and compact). Dust model dependent on stellar models?
\end{itemize}

@@

Under these assumptions, we would like to infer the true parallax $\varpi_{\true}$ for each star given its measured parallax $\varpi$ and associated uncertainty $\sigma_{\varpi}$ from \tgas.
The strength of our inference method here lies in how we generate our prior information on the true parallax.
Supplementary to this trigonometric measure of distance, the photometry of stars is also informative of their distances.
The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the color--magnitude diagram (\cmd).
The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star, when compared with its brightness or apparent magnitude, contains distance information.
We include this distance information in the inference of the true parallax by using the \cmd\ as our prior.
Specifically, we use the \cmd\ generated from the data themselves as our prior.
Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
This \cmd\ model becomes, for each star---given its observed color and magnitude (and uncertainties on those)---a prior pdf for the distance to that star.

@@

\begin{equation}
10^{0.2\,M^c_{J,n} + 2} = \varpi\,10^{0.2J^c_n}
\label{eq:transform}
\end{equation}

@@

If life were simple, our inference would be a 1D problem using Bayes theorem and look something like:
\begin{equation}
p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
\label{eq:bayes}
\end{equation}
where $p(\varpi_{\true})$ is a prior probability density (pdf) for the
parallax to the star, and $Z$ is a normalization constant.

But life isn't simple, and our prior is a 2D function in color and transformed absolute magnitude space, so our full inference in done in 2D space as well.
We assume (as \gaia\ suggests) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood is
\begin{equation}
p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) = \mathcal{N}(\mathbf{y}_n \given \mathbf{Y}_n, V_n)
\end{equation}

@@

where $\mathcal{N}$ is a 2D Gaussian function with mean $\mathbf{Y}$ and variance $V_n$.
Here we have assumed that the parallax uncertainty dominates over the $J$ band uncertainty, so we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty. This is to keep our uncertainties Gaussian so we can use \xd\ generate our prior below.

@@

To generate our prior, we fit the data in the slightly transformed \cmd\ space, Equation \ref{eq:transform}, using \xd\ which deconvolves the data with its uncertainties, which are heterogeneous measurement uncertainties. It generates the underlying distribution from which the uncertain data were drawn from. It represents the distribution as a mixture of Gaussians which maximizes the marginalized likelihood of the data, marginalized over the true data.

@@

We use stars crossmatched in TGAS, 2MASS and within the observing footprint of \panstarrs\ to access the Greene 3D dust model. We require the photometry have real values and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.

@@

We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from iteratively sampling our posterior distances. Dust corrections are non trivial for stars with poor signal to noise. Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections. This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes. To apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry. Using this raw prior we infer more precise distances to all the stars. We then use this more precise distance posterior to get a measurement of dust in the 3D dust map. This approach minimizes severe distances from the likelihoods alone and their associated severe dust corrections. In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance. We take the $5\%$ quantile, as apposed to the median for example, to do a light dust correction because over correcting for dust can shoot the likelihood into an unphysical, different space of the CMD and then bias the distance inference. For each distance we sample the median of the probabilistic dust model. We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models. We then regenerate our prior using the lightly dust corrected photometry. We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution. We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept. We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations. The iterative process allows for slight variations in distance and corresponding dust to become more consistent with the \cmd.

@@

As proof of concept we inferred the distances to stars in the open cluster M67. It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away. Previous work has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5. With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, for stars with the most certain parallaxes, the SN reaches 5 at about 700 pc, right where this cluster lies. So it is a great test case for comparing parallax inferences with or without a prior. We made a window of 1 degree squared on the sky with the cluster M67 lying in the middle. The cluster is visibly obvious as a clustering of stars in a visualization of the \gaia\ data in 2D on the sky. With our converged dust values, and it's associated prior, we calculate the posterior of the true parallaxes to the stars within this window on the sky around M67. Figure XXX shows the comparison of the likelihoods and posteriors. The vertical lines bracket the parallaxes of, or distances to, the stars associated with the cluster. The likelihoods of the observed parallaxes are far broader and the cluster is not obvious at all. In the posteriors there is a sharp peak in the probability at the previously determined distance to the cluster, the vertical lines, for far more stars than the likelihoods would imply. Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy. Too strong?

@@

Come back to the assumptions. What happens if we change or violate them?

>>>>

In this paper we use the noise deconvolved observed \cmd\ as a prior to infer more precise parallax pdfs from \gaia\ measurements.
Understanding this method is challenging because we are using the data both for the prior and the likelihood in our inference.
To ease understanding, here we will demonstrate this method with a simpler toy model.
Instead of inferring more precise parallax pdfs using the 2D distribution of the \cmd\ as our prior, for this simpler model we will infer more precise pdfs of the value $Y$.
We will do this simpler inference using a 2D distribution that is a 1D Gaussian in the y-direction, with a running mean $\mu = mx + b$ and some thickness $t$, as our prior.
Here the thickness of the distribution mimics the true thickness of the \cmd\ due to age, mass and metallicity.
The $1\sigma$ contours as well as (noise free) samples from the toy model distribution are shown in Figure \ref{fig:toy}, upper left panel.
Similarly, noisy samples from this distribution are shown in the upper right panel.
Using a representation of the true distribution (which we learn from the noisy data) as our prior, we can infer some posterior belief of $y_{\true,n}$, a true sample of the distribution, from $y_n$, a noisy sample of the distribution.

Setting up the problem in more detail, we have the true underlying distribution of points and the noisy data

\begin{eqnarray}
y_{\true, n} \, &=& \, m_{\true}\,x_n + b_{\true} + \Delta_n \\
y_n \, &=& \, y_{\true,n} + \epsilon_n
\label{eq:ytrue}
\end{eqnarray}

where $\Delta_n$ is drawn from the Gaussian function $\mathcal{N}(\Delta_n \given 0, t^2)$, and $\epsilon_n$ is drawn from the Gaussian function $\mathcal{N}(\epsilon_n \given 0, \sigma_n^2)$, so $\sigma_n$ is the measurement uncertainty.
These represent true and noisy draws from a distribution that is a straight line with some thickness $t$, and are visualized in the upper panels of Figure \ref{fig:toy}
The upper left shows 1024 realizations of the true values, and the upper right, as well as lower left, show 1024 realizations of the noisy values $y_n$ and their associated measurement uncertainties $\sigma_n$ as the grey points. Here we highlight 5 random points in black to highlight the method, but the same method is being applied to all the data.

To infer a pdf of the true y value for each measured y value we need a prior. For our prior we will use an estimate of the underlying true distribution that the noisy data $y_n$ are drawn from, which we build from the noisy y data themselves.
We learn the noise-deconvolved function using marginalized maximum likelihood estimation (MML), which marginalizes over the true data values (we never need to know them) and maximizes the likelihood of the data. Here our likelihood for each individual datum is

@@

Here is a good point to step back and acknowledge that our method is odd/strange/different/new (which maybe you're also currently thinking but can't quite express why yet). We are using the set of data to determine the noise-deconvolved distribution, and then using that noise-deconvolved distribution to make inferences on the data. We are technically using the data twice! If we were going to be fully, technically correct, we should learn the distribution, but leave one datum out; the one datum we want to infer the true y value for\footnote{DWH: Remember what this brain dump is for}. Leaving one datum out to estimate the prior and then inferring the posterior for that datum is way too expensive for the 1024 points in our toy model, and also for the 1.2 million stars in our actual data set. Luckily with large data sets, even at 1024, 1 datum is $\sim0.1\%$ of your dataset and therefore leaving that point out will not change the MML parameters of your noise-deconvolved distribution, and with the \gaia\ dataset of over a million stars, we can safely step into the Empirical Bayes regime. (Talk more about emperical bayes here?)

@@

Since everything is Gaussian, this can be done analytically by multiplying Gaussians together. In Figure \ref{fig:toy} lower left panel, the blue points represent the posterior belief of the true value $p(y_{\true,n} \given y_n, \tilde{\sigma}_n)$ for the highlighted black points. The red points are the true y values.
The blue points in Figure \ref{fig:toy}, lower right panel, show the expectation value and $1\sigma$ uncertainty of the posterior belief for all 1024 $y_n$.
When the true values are inferred in this way, the associated variance $\tilde{\sigma}_n^2$ is the convolution of the true distribution variance $t^2$ and the measurement variance $\sigma_n^2$.
The points are colored by their measurement variance $\sigma_n^2$ which shows that noisier measurements are more influenced by the prior and get pulled closer to the center of the MML distribution, our prior. In contrast, more certain measurements remain closer to their measured values $y_n$.
By inferring the true y value from the measured y value in addition to the prior information of the distribution the y value was drawn from, this we have better estimates of the true y value for each point. Bringing the toy model back to our original problem, using the noise-deconvolved CMD as our prior allows us to get better estimates of the parallax for each star.

@@

We make the following assumptions. The method we present is correct
and justifiable under these assumptions, each of which is questionable
in its own right. We return to criticize these assumptions---and the
method that flows from them---in the Discussion below.
\begin{description}
\item[stationarity] We make an assumption of \emph{stationarity}; that is, that the
  stars measured at lower signal-to-noise have analogs measured at
  higher signal-to-noise.  In detail, this assumption is wrong, of
  course (for example, distant halo stars are different than local
  disk stars), but the stationarity assumption is fairly weak; as long
  as there is support in the high signal-to-noise stars for the types
  of stars seen at low signal-to-noise, the results will not be
  strongly biased, but this assumption is deep and fundamental to this
  work.
\item[big data] We assume that we have large numbers of stars, large enough that
  an empirical-Bayes or maximum-marginalized likelihood is a safe
  approximation to full Bayesian inference. This assumption will be
  least true in the least-populated parts of the \cmd. In the extreme
  case of a stellar type with only one (statistically distinct)
  exemplar in the data set (for example, a white dwarf), this assumption will be maximally wrong.
\item[noise model] We assume that the \tgas\ parallax uncertainties dominate the
  noise in any estimates of absolute magnitude. We further assume that
  the parallax and color uncertainties are Gaussian in form with
  correctly known variances. In detail, we assume that the [INSERT
    CORRECT TERMINOLOGY FROM GAIA HERE] parallax uncertainty estimates
  are correct. Here we are making use of the fact that, for
  de-noising, under-estimation of measurement noise is conservative
  (it leads to under-deconvolution).
\item[dust] We treat the \panstarrs-based three-dimensional dust maps (\citealt{green15})
  as correct in their median dust estimates, and their
  effects on color and magnitude. As we optimize the empirical-Bayes model, we iteratively update each star's
  dust correction at its updated, inferred distance. Our assumption is that these corrections are
  good enough, and uncertainties in these are not dominant, for either
  color or absolute magnitude.
\item[mixture of Gaussians] We assume that the \cmd\ can be represented with a mixture of
  Gaussians in a particular transformed space. This mixture we fix
  (with only heuristic analysis) at 128, 512, and 2048 Gaussians in
  different experiments.
\item[no physics] We make no use of stellar models; our only assumptions about
  stellar physics are generic and implicit (for example, that the
  \cmd\ is somehow smooth and compact).
\end{description}

@@

Under these assumptions, we would like to infer the true parallax $\varpi_{\true}$ for each star given its measured parallax $\varpi$ and associated uncertainty $\sigma_{\varpi}$ from \tgas.
The strength of our inference method here lies in how we generate our prior information on the true parallax.
Supplementary to the trigonometric measure of distance, the photometry of stars is also informative of their distances.
The intrinsic temperatures and luminosities of stars tightly correlate leading to the observed colors and absolute magnitudes of stars also tightly correlating, usually visualized as the color--magnitude diagram (\cmd).
The tight correlation of the \cmd\ benefits our inference because the luminosity or absolute magnitude of a star, when compared with its brightness or apparent magnitude, contains distance information.
We include this distance information in the inference of the true parallax by using the \cmd\ as our prior.
Specifically, we use the \cmd\ generated from the data themselves as our prior.
Instead of assuming a functional form for the spatial distribution of stars in the Milky Way (\citealt{astraatmadja16b}), or relying on stellar models (\citealt{gaia16}), we build a noise-deconvolved estimate of the \cmd\ from the observed colors and absolute magnitudes of our full set of stars.
This \cmd\ model becomes, for each star---given its observed color and magnitude (and uncertainties on those)---a prior pdf for the distance to that star.

@@

\begin{equation}
y_0 = \varpi\,10^{0.2J^c_n} = 10^{0.2\,M^c_{J,n} + 2}
\label{eq:transform}
\end{equation}

@@

If life were simple, our inference would be a 1D problem using Bayes theorem and look something like:
\begin{equation}
p(\varpi_{\true} \given \varpi, \sigma^2_{\varpi}) = \frac{1}{Z}\,p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi}) \, p(\varpi_{\true}) \quad ,
\label{eq:bayes}
\end{equation}
where $p(\varpi \given \varpi_{\true}, \sigma^2_{\varpi})$ is the likelihood of the observed parallax, $p(\varpi_{\true})$ is a prior probability density (pdf) for the
parallax, and $Z$ is a normalization constant.

But life isn't simple, and our prior is a 2D function in color and transformed absolute magnitude space, so our full inference in done in 2D space as well.
We assume (as \gaia\ suggests) that the parallax likelihood is Gaussian, as well as the $(J-K)^c_n$ color likelihood. So our full likelihood of the data is
\begin{equation}
p(\mathbf{y}_n \given \mathbf{Y}_n, I_n) = \mathcal{N}(\mathbf{y}_n \given \mathbf{Y}_n, V_n)
\end{equation}

@@

where $\mathcal{N}$ is a 2D Gaussian function with mean $\mathbf{Y}$ and variance $V_n$.
Here we have assumed that the parallax uncertainty is significantly larger than the $J$ band uncertainty, and therefore the dominant uncertainty in the transformed absolute magnitude, so we take $J_n^C$ as a point estimate and only propagate the parallax uncertainty. This is to keep our uncertainties Gaussian so we can use \xd\ generate our prior below.

@@

To generate our prior, we fit the data in the slightly transformed \cmd\ space, using Equation \ref{eq:transform} instead of absolute magnitude. We fit this transfromed \cmd\ using \xd\ which deconvolves the data with its uncertainties, which are heterogeneous measurement uncertainties. It generates the underlying distribution from which the uncertain data were drawn from. \xd\ generates a represention of the distribution as a mixture of Gaussians which maximizes the marginalized likelihood of the data, marginalizing over the true data.

@@

We use stars crossmatched in \tgas\ and \tmass\, and lie within the observing footprint of \panstarrs\ to access the Greene 3D dust model. We require the photometry have real values, and nonzero, real errors, and remove a small selection of \tmass\ stars that have zero color and zero J band magnitude.

@@

We correct for dust using 3D Bayestar (Schlafly+Finkbeiner) with an estimate of the distance from iteratively sampling our posterior distances. Dust corrections are non trivial for stars with poor signal to noise. Stars with poor signal to noise have a likelihood that is consistent with infinite distance and can therefore have severe dust corrections. This is most obvious in the giant stars, which tend to have the lowest signal to noise, and can therefore have dust corrections $> 2$ magnitudes. To apply a dust correction to the \tmass\ photometry, we first generate our prior using the observed, attenuated photometry. Using this raw prior we infer more precise distances to all the stars in our sample. We then use this more precise distance posterior to get a measurement of dust in the 3D dust map. This approach minimizes severe distances from the likelihoods alone and their associated severe dust corrections. In detail, we take the $5\%$ quantile of each distance posterior, so the closest bit of the posterior, and sample the Green et al 3D dust map at that distance. We take the $5\%$ quantile, as apposed to the median for example, to do a light dust correction because over correcting for dust can shoot the likelihood into an unphysical, different space of the CMD and then bias the distance inference. For each distance we sample the median of the probabilistic dust model. We apply each light dust correction to our \tmass\ photometry given by Equation \ref{eq:data} where $Q_{\lambda}$ = [0.709, 0.302] for bands [J, K] respectively, taken from table (REF TABLE SFD), and $E(B-V)$ is the output of the Green et al dust models. We then regenerate our prior using the lightly dust corrected photometry. We do not update the errors, but keep them the same as before the dust correction, which is wrong and under representing our errors but this leads again to a lighter deconvolution. We could add the covariance of the dust to our likelihoods, but this is beyond the scope of this demonstration of concept. We iterate this process 10 times, however the dust values seem to converge after about 2-3 iterations. The iterative process allows for slight variations in distance and corresponding dust to become more consistent with the \cmd.

@@

As proof of concept we inferred the distances to stars in the open cluster M67. It is an open cluster in the constellation Cancer and is estimated to be 3-5 billion years old and 800-900 pc away. Previous work (cite CBJ) has estimated that bad things happen if you try to infer a distance as the inverse of the observed (noisy) parallax when the signal to noise of the parallax measurement drops below 5. With \gaia\ parallaxes having a minimum uncertainty of about 0.3 mas, the SN reaches 5 for the most certain parallax measurements at about 700 pc, right where this cluster lies. So it is a great test case for comparing parallax inferences with or without a prior. We made a window of 1 square degree on the sky with the cluster M67 lying in the middle. The cluster is visibly obvious as a clustering of stars in a visualization of the \gaia\ data on the 2D sky. With our converged dust values, and it's associated prior, we calculate the posterior of the true parallaxes to the stars within this window on the sky around M67. Figure \ref{fig:m67} shows the comparison of the likelihoods and posteriors. The vertical lines bracket the parallaxes of, or distances to, the cluster. The likelihoods of the observed parallaxes, shown in the top row, are broad and the cluster is not obvious at all. The posteriors, shown in the bottom row, show a sharp peak in the probability at the previously determined distance to the cluster, the vertical lines, for far more stars than the likelihoods would imply. Our prior is not only making parallaxes more precise but also (possibly) increasing the accuracy. Too strong?

@@

Come back to the assumptions. What happens if we change or violate them? Long discussion of the noise model. We took serious short-cuts. Call out to Boris.

LMA: Some discussion of the inconsistency of how we treat dust within the probabilistic framework.
=======
Come back to the assumptions. What happens if we change or violate them?
